---
title: "pearsons_correlations"
author: "Juliana Cornett"
date: "2023-10-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Pearson's correlations

#read in data

```{r}
oyster_psts <- read.csv("../Data/Oyster_PSTs.csv")
alex <- read.csv("../Data/Alexandrium_MicroscopeCounts.csv")
```

#PHYTOPLANKTON

#Pearson's correlation between collection methods

#calculate Pearson's correlation between 1m and 5m counts

```{r}
correlation_1m_5m <- cor.test(alex$Count_1m_L, alex$Count_5m_L, method = "pearson")
correlation_1m_5m
```

#visualize relationship between 1m and 5m counts

```{r}
plot_1m_5m <- ggplot(alex, aes(Count_1m_L, Count_5m_L)) + geom_point() + geom_smooth(method = "lm") + theme_pubr()
plot_1m_5m
```

#calculate Pearson's correlation between 1m and net tow

```{r}
correlation_1m_net <- cor.test(alex$Count_1m_L, alex$Count_Net, method = "pearson")
correlation_1m_net
```

#visualize relationship between 1m and net tow

```{r}
plot_1m_net <- ggplot(alex, aes(Count_1m_L, Count_Net)) + geom_point() + geom_smooth(method = "lm") + theme_pubr()
plot_1m_net
```

#calculate Pearson's correlation between 5m and net tow

```{r}
correlation_5m_net <- cor.test(alex$Count_5m_L, alex$Count_Net, method = "pearson")
correlation_5m_net
```

#visualize relationship between 5m and net tow

```{r}
plot_5m_net <- ggplot(alex, aes(Count_5m_L, Count_Net)) + geom_point() + geom_smooth(method = "lm") + theme_pubr()
plot_5m_net
```

#TOXINS