---
title: "psts_plot"
author: "Juliana Cornett"
date: "2023-08-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#code to generate plot of PST levels over the course of the summer (Figure 4 in manuscript?)

#install packages (ggplot2 & ggpubr for creating figure)

```{r}
#install.packages("ggplot2")
#install.packages("ggpubr")
```

#load libraries

```{r}
library(ggplot2)
library(ggpubr)
```

#read in ELISA PSTs data

```{r}
oyster_psts <- read.csv("../Data/Oyster_PSP_Data.csv")
```

#put date in date format for timeseries

```{r}
oyster_psts$Collection_Date <- as.Date(oyster_psts$Collection_Date, "%m/%d/%Y")
```

#select summer 2023 data only (5/4/23 and later)

```{r}
psts_summer23 <- subset(oyster_psts, Collection_Date >= "2023-05-04")
```

#timeseries plot (maybe add colored shading for when farm was shut down/testing above DEC levels? Or add in DEC levels in a different color?)

```{r}
pst_plot <- ggplot(psts_summer23, aes(x = Collection_Date, y = Tox_ug_100g)) +
  geom_point(size = 3) +
  theme_pubr() +
  scale_x_date(date_breaks = "1 month", date_labels = "%b %Y") +
  labs(x = NULL, y = "PSTs in Oyster Tissue (ug/100g)") 
pst_plot
```

#save as jpeg

```{r}
jpeg("../Figures/Fig4.jpg", width = 8, height = 5, units="in", res=600)
pst_plot
```
